---
title: "Nobel-vignette"
author: "Zijin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, echo = FALSE}
library(nobel)
library(dplyr)
library(ggplot2)
library(purrr)
library(tibble)
```

# get data from API
The two basic functions get data from the laureate endpoint and prize endpoint seperately. They do not need any parameter, just use them to get the whole dataset.
```{r}
head(get_laureate(),5)
head(get_prize(),5)
```

# Analyze with the package nobel
## get info
Simply get basic information with category and year information.
```{r}
get_nobel_info(c = "literature", y = "2012")
```

## age
This function shows the age when laureate winning the prize, select the data with the year or category or both, or directly use the full name of the winner.

Here is an example of geting the age information of winners in the year 2000.
```{r}
get_nobel_age(y = "2000")
```

See the trend of the age. Make it into a graph.
```{r}
compare_age <- get_nobel_age(c = "physics")%>%
  select(category,year,age)
```
```{r message = FALSE}
ggplot(data = compare_age, aes(x = year, y = age, group = category,color = category))+
  scale_x_discrete(breaks=seq(1901,2020,20))+
  geom_smooth()
```

## gender
This function shows the table of gender or orgnization given the year ot category. Users can choose whether includes the orgnization information in the output besides input the year or category.

Here is an example of getting gender information of the year 1997.
```{r}
get_nobel_gender(y = "1997")
```

Calculate the ratio of female/male in different category.Then make it into a barplot to compare the female/male ratio of winning among categories. 
```{r}
gender <- list()
for(c in c("physics","peace","economics","medicine","chemistry","literature")){
  gender[[c]] <- get_nobel_gender(c = c, org = FALSE)
}
```
```{r}
female <- map(gender,"n")%>%
  map(as.numeric)%>%
  map_dbl(1)%>%
  as.data.frame()
colnames(female)<-c("female")
male <- map(gender,"n")%>%
  map(as.numeric)%>%
  map_dbl(2)%>%
  as.data.frame()
colnames(male)<-c("male")

ratio <- cbind(female,male)%>%
  rownames_to_column(var = "category")%>%
  mutate(ratio = round((female/male)*100,2))%>%
  select(category,ratio)%>%
  arrange(ratio)
ratio
```

```{r}
ggplot(data = ratio, aes(x=category,y=ratio))+
  geom_bar(stat="identity",fill="steelblue",width=0.5)
```

## country
This function shows the rank of countries(only shows the name of the existed country) where winners were born given a year or category.

Here is an example of country rank in year 1997.
```{r}
get_nobel_country(y = "1997")
```

The number of winning of the U.S every year from 2000-2010.
```{r}
country <- list()
for(y in c(2000:2010)){
  y <- as.character(y)
  country[[y]] <- get_nobel_country(y=y)%>%
    filter(born_country == "USA")
}
```
```{r}
usa <- map_dbl(country,"n")%>%
  as.data.frame()%>%
  rownames_to_column()
colnames(usa) <- c("year","winner")

ggplot(usa, aes(x = year, y = winner, group = 1))+
  geom_line()
```

## share
This function shows the share of the prize by given year or category or both.

Here is an example of share information of the prize in 2019.
```{r}
get_nobel_share(y = "2019")
```

Which category is most likely to share the prize with more winners? Get the average share of different category over 1901-2019.
```{r}
get_nobel_share()%>%
  group_by(category)%>%
  summarize(share = round(mean(share),2))%>%
  arrange(desc(share))
```
